---
title: "Nested SDM Script for Island Plants - AKA OCN 750 HW 11"
author: "M Kapur"
date: "May 25, 2016"
output: html_notebook
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```
####Consider that the data are naturally organized into groups (site/island/archipelago/region), and we should account for this structure in the model. In addition, some of the predictors are defined at the scale of the archipelago, and so what they could predict is archipelago-mean richness. We may have to establish at which scale each predictor is defined before throwing them all into a large model.
```{r, warning = FALSE, message=FALSE}
require(lme4) 
require(MuMIn)
require(lattice)
require(effects)
require(lmerTest)

isle = read.csv("cabral_island_data.csv")
arch = read.csv("cabral_arch_data.csv")
```

#### Relatively how much variation in richness is there, between vs. within archipelagos? Which archipelagoes are particularly diverse, and which are depauperate?

```{r}
mod1 <- lmer(Species ~ (1|Archipelago), data = isle)
## The typical difference across Archipelagos is 164.5, and within is 147.8. So there is not a massive difference between islands vs. within.


## this function identifies which Archipelago deviates the most positively, which is Canaries.
## Once we have different levels within the model we can check those as well.
max(ranef(mod1)$Archipelago)

```

#### Exploratory plots of the effect of each variable on richness, plotted at the appropriate scale. (Datasets merged). Some predictors might need to be transformed for use in a linear model.

```{r, echo=FALSE}
div = merge(arch, isle, by = "Archipelago")
nombre = names(div)
nombre = nombre[-c(1,2,6,7)]

par(mfrow = c(2,3))
for (i in nombre){
  plot(div$Species, div[,i], data = div, ylab = i, xlab = "spp richness", main = paste0(i, " vs. Richness"), pch = 19)
}
dev.off()
plot(Species ~ Archipelago, data = div)
```
#### A mixed model that includes all the predictors. Do the usual rigmarole: plot fitted effects (fixed and random), plus model diagnostics. How much variation do the predictors explain, at the two scales in the data (island and archipelago) & how much of the total variation have they explained, according to R2GLMM(m)?
```{r, warning=FALSE,message=FALSE}
## this is the model with all env predictors
mod2 = lmer(Species ~ (1|Archipelago) + number.islands + distance + age + Area + Elev + Temp, data = div) ## Archipelago is higher-level predictor
VarCorr(mod2)


## we could then update the model with other ranef parameters like site, island, region etc:
#update(mod1, . ~ . + (1|Archipelago) )

par(mfrow = c(1,2))
## is normality assumption met?
qqmath(ranef(mod2))
#dev.off()
plot(resid(mod2) ~ fitted(mod2), pch = 19, main = "residuals vs fitted values") 
## looks good

require(effects)
plot(allEffects(mod2))

## R2 GLMM
require(lattice)
dotplot(ranef(mod2))

require(MuMIn)
r.squaredGLMM(mod2)

```

####Use F-tests to test the significance of the predictors -- e.g. what is most important in controlling species richness in islands and archipelagos? What are the denominator degrees of freedom for each predictor? This is essentially telling you how much replication there is for that predictor, minus the number of parameters for that predictor. 
```{r}
require(lmerTest)
anova(mod2, ddf = "Kenward-Roger") 
## All values indicate 1 df. 

```
####Presumably you found that distance from the mainland is an important predictor. Use 1) the fixed effect estimate for distance, and 2) the random effect estimate for archipelago-mean richness, to plot the model prediction for each archipelago vs. distance from the mainland for that archipelago. This is like a partial residual plot at the level of archipelagos, which will give you a sense for how well distance predicts archipelago-mean richness.
```{r}
##fixed effect for distance
fixef(mod2)[3]

##random effect for archipelago richness
rand.eff = ranef(mod2)$Archipelago[,1] + fixef(mod2)[1]

## plot model prediciton vs distance
plot(rand.eff[1:23] ~ arch$distance[1:23], pch = 19, xlab = "distance from mainland", ylab = "predicted value", main = "distance vs predicted richness")

```
####Is the model the best model? Often I just stick with one big model when the ratio of data to parameters is pretty good. But some people really like to find the best model, or to look at model weights, etc. Use AICc in some capacity to assess which predictors are important, what the best model is, and how sure you are about what the best model is. Remember to do REML=FALSE for comparing models.
```{r}
AICc(mod1, REML = FALSE)
AICc(mod2, REML = FALSE)
model.sel(c(mod1,mod2), REML= FALSE)

## the second model (with more parameters) is much better supported. Model 1 only includes Archipelago, so there isn't an ability to compare important parameters there.
```

