---
title: "OCN 750 - HW14"
author: "Maia Kapur"
date: "Wednesday, December 02, 2015"
output: html_notebook
---

#### To start off, see how the environmental parameters covary with each other. Do a PCA on the environmental predictors. To make the results more meaningful, you'll probably want to log- or square-root-transform some of the predictors, because some of them are highly skewed. How much variation is explained by the PC axes? Are there dominant axes of variation, or is the environmental variation more complex than that? Make a biplot to look at how the variables load onto the first two axes. Also make a biplot for the third and fourth axes.
```{r, warning = FALSE, message = FALSE}
require(vegan)

macro = read.csv("macroinverts.csv")
macro$Cond = with(macro, log(macro$Cond)) ## log transform skewed Cond predictor


macro.sub = subset(macro, select = names(macro[2:12])) ## subset data to environmental predictors
macro.pca = rda(na.omit(macro.sub), scale = TRUE) ## omit NAs and run PCA

## plot it
biplot(macro.pca, choices = c(1,2), main = "PCA, Ords 1 & 2")
biplot(macro.pca, choices = c(3,4), main = "PCA, Ords 3 & 4")

## it seems to be a little complex -- there appears to be a slight clustering of large - to small substrate types (e.g. Gravel vs Bedrock)
```
#### Eventually you're going to do a constrained ordination, using environmental variables to predict invertebrate community structure. But right now there are just too many predictors (and I didn't even include all those the authors looked at). Reduce the number of variables in the environmental data to a maximum of 6. Picking variables to exclude is a judgment call, but if two variables seems pretty strongly correlated with each other, then dropping one of them makes sense. Or you can use envfit() or capscale() to drop variables that don't seem important for community composition. It's up to you.

```{r, echo=FALSE}
macro.sub2 = subset(macro, select = !(names(macro)) %in% c("T", "PO4", "Boulder", "Cobble", "Gravel", "Sand", "Clay.silt", "Pebble")) ## drop some environmental variables
```
#### Now use NMDS to ordinate the invertebrate community data. Some of the families
are quite rare, so drop those that occur less than 10 times. You should also
transform the count data before using it to calculate dissimilarities. I've been using
the wisconsin square root transformation in class, but it's not the only option. 
```{r}
## a for loop to identify which columns have less than 10 non-zero occurences (ie are rare), then passes the names of these columns to a vector called "drops".
drops = NULL
for (n in 8:82){
  if (sum(macro.sub2[n] != 0) < 10){
    drops <- append(drops, names(macro.sub2[n]))
  }
}
macro.sub3 = subset(macro.sub2, select = !names(macro.sub2) %in% drops) ## subsets the dataframe for columns NOT included in the drops list

macro.sub3 = wisconsin(sqrt(na.omit(macro.sub3[2:33])))

```
####Try NMDS with 2-4 dimensions. How much does the stress change with the # of dimensions?
```{r, message = F}
ord2 = metaMDS(macro.sub3, dist = 'bray', k = 2)
ord3 = metaMDS(macro.sub3, dist = 'bray', k = 3)
ord4 = metaMDS(macro.sub3, dist = 'bray', k = 4)
## stress decreases from ~ 0.2 to ~ 0.1 with increasing dimensionality
```
####Perform a principal coordinates analysis (with capscale), and compare the ordination to the NMDS ordination using procrustes(). Do the two methods give similar results? How much variation in community composition is explained by the first few PCoA axes?
```{r}
pcoa = capscale(macro.sub3 ~ 1, dist = 'bray')
head(summary(pcoa)) ## about 50% of variation explained by first four components
proc = procrustes(pcoa, ord4)
plot(proc)
```
####Using the subset of environmental variables you chose earlier, use envfit() to see how the environmental variables relate to the NMDS ordination axes. Plot the results. Which variables seem most strongly related to the major axes of community variation?
```{r}
fit2 = envfit(ord2 ~ pH + Cond + oxygen + Turb + Bedrock  + tmax, data = macro.sub3, na.rm = T)
fit4 = envfit(ord4~ pH + Cond + oxygen + Turb + Bedrock  + tmax, data = macro.sub3, na.rm = T)

ordiplot(ord2, display = "sites", type = 'n', main = "nmds 2d")
points(ord2)
plot(fit2, col = "seagreen3", arrow.mul = 0.7)


ordiplot(ord4, display = "sites", type = 'n', main = "nmds 4d")
points(ord4)
plot(fit4, col = "purple2", arrow.mul = 0.7)
## Looks like Bedrock and Turbidity are the strongest related to NMDS2, with Cond less-strongly associated with NMDS1.

```
####Now use constrained ordination (canonical analysis of principal coordinates) to more directly look for axes of community variation that are explained by the environment. Plot the site scores and the biplot arrows for the environmental axes. How are the environmental variables aligned along the most important axes? How much variation in community composition is explained by those axes? 
```{r}
cap = capscale(macro.sub3 ~ pH + Cond + oxygen + Turb + Bedrock + tmax, data = macro.sub3, dist = "bray")
head(summary(cap))
## the first two constrained components explain  66% of 56% of the variation, or 36% of the overall variation
plot(cap, display = "sites", type = "n")
points(cap, display = "sites", col = "lavenderblush3")
text(cap, display = "bp", col = "cadetblue4")
## none of the variables seem perfectly correlated with the axes, but it appears that chemical indicators vary positively with CAP1, while less-fine benthic structure (e.g. pebbles) vary positively with CAP2.
```
####Test whether the axes are 'significant', using the permutation tests provided by the anova() function. Use a plot of the species scores, in combination with the environmental vectors, to see which invertebrate families respond most strongly to the different environmental axes.
```{r}
anova(cap, by = "margin") 

## Cond, Turb, Bedrock and tmax are significant
plot(cap, display = 'species')
text(cap, display = 'bp', col = "darkgrey")
## No species vary positively with salinity, though some like Oligochaeta, Physidae and Lymnaeidae seem to correlate with tmax somewhat. The majority are negatively correlated with both.
par(mfrow = c(1,2))
plot(macro.sub3$Oligochaeta, macro.sub3$tmax, main = "Oligochaeta vs. Tmax")
plot(macro.sub3$Oligochaeta, macro.sub3$Cond, main = "Oligochaeta vs. Salinity")

```
You've been looking at overall variation in community composition, but the
authors of the study were most interested in the effects of salinity and pesticides.
Use permanova (i.e., adonis()) to test the effect of these predictors on community
composition. Are there significant effects? How much variation is explained? Is
there an interaction between the two predictors? Would you consider these to be
strong environmental impacts, at least in terms of changing the structure of the
stream invertebrate communities?
```{r}
adonis(macro.sub3 ~ tmax + Cond, data = macro.sub3, dist = "bray" )
## automates the marginal test (leave-one-out), caveat is it does it in order of appearance
## There appear to be significant effects for both predictors.
adonis(macro.sub3 ~ tmax + Cond + tmax*Cond, data = macro.sub3, dist = "bray" )
## interaction is not significant
## Based on the capscale, they do appear to be significant drivers both visually and quantitatively, but the direction doesn't seem so much as to favor different regimes, more just to inhibit all except a few species. 
```

