---
title: "OCN 750 HW 12 - Revisiting Heat Shock"
author: "Maia Kapur"
date: "Thursday, November 12, 2015"
output:
  html_document: default
  html_notebook: default
---

#### Fit a GLMM that incorporates random effects for isofemale and population, and also includes fixed effects and interactions that test the 6 questions. We also need to account for potential overdispersion in the binomial response. The best way to do this in lme4 is with an 'individual level random effect'. The thing to do is make a new factor where each row of the dataset gets a level. This is done in the replicateID column which is just a single number as a factor that gets popped into the dataframe. Other ways of doing this include the seq() function.

```{r, message=FALSE, warning=FALSE}
require(lme4)
require(car)
require(effects)
require(lattice)
heat = read.csv("heat_shock_all.csv")
heat$replicateID = factor(1:nrow(heat))
#heat$proportion[heat$proportion != 0] <- 1 ## change any positive proportion of survivorship to 1
mod1  = glmer(proportion  ~  (1|isofemale) + (1|population) + sex + region + treatment + treatment*region + treatment*sex + sex*region + (1|replicateID),  data	= heat, family = "binomial")

```
#### Use likelihood ratio tests to see which of the interactions is significant (LRTs tend to be anti-conservative for fixed effects, but this model has a lot of data, and doing a parametric bootstrap would take a long time with this model). Drop the non-significant interactions, which will allow you to test and interpret the main effects more clearly. Now use LRTs to test the significance of the remaining fixed effects in the model. <span style = "color:red">MK note on 11/7/16: these are taking a weirdly long time to run. </span><span style = "color:blue">MK note on 11/14/16: the first upload of this didn't explicitly drop ALL the non-significant terms identified in mod1. Now I've only included sex and treatment in add'n to the random effects </span>
```{r, message=FALSE, warning=FALSE}
summary(mod1)
## Only intercept (sexFemales) and the treatment are significant (treatment only marginally so.) New model drops all interaction terms and includes random effects.
mod2  = glmer(proportion  ~  (1|isofemale)  + (1|population) + (1|replicateID) + treatment + sex,  data  = heat, family = "binomial")
summary(mod2)
anova(mod1, mod2, test = "Chisq")

## The second model has a slighly lower AIC value.
```

####Plot the fitted fixed effects using the effects package. How do you interpret these results, in light of the 6 questions listed above? <span style = "color:red">MK note on 11/7/16: allEffects plots sometimes get squished. I coerce these to be big enough by using fig.width = and fig.height = in the brackets at the code header. </span>
```{r, message=FALSE, warning=FALSE, fig.width = 9, fig.height = 9}
plot(allEffects(mod1))
plot(allEffects(mod2))
## It appears from the effects plot of the full model (mod1) that the only significant differences are between SW Hardening and Caribbean Base (eg treatment * region) and differences between treatments alone but not between sexes and treatments. The effects plots for mod2 suggest treatment is the most predictive.
     
```
#### Use the random effects estimates, and dotplot() of those estimates, to discuss how much genetic variation there is among populations, and among isofemales within populations, and how much extra (overdispersed) variation there is among replicates. Plotting the individual-level random effect for replicate is not advised; there are too many levels for the function. How does the spread for the random effects compare to the effect sizes of the fixed effects?
```{r, message=FALSE, warning=FALSE}
par(mfrow = c(2,2))
#dotplot(ranef(mod2))
fixef(mod2)  

```
####The authors were interested in another question as well: Does acclimation to high temperature (i.e., the treatment effect) exhibit genetic variation among isofemale lines? To test this, you need to allow treatment to vary by isofemale. In other words, you need to include a fixed*random interaction. Put this in the model, and use a likelihood ratio test to see if it is important. Use dotplot() to visualize how the treatment effect varies by isofemale. Does the genetic variation seem substantial? What is the correlation between the random Intercept and random treatment effect among isofemales? Can you explain what this correlation implies? Does including the treatment-by-isofemale interaction change the results for the fixed effects in the model? If so, how?
```{r, message=FALSE, warning=FALSE, fig.width = 12, fig.height = 9}
mod3  = glmer(proportion  ~  (1|isofemale)  + (1|population) + (1|replicateID) + treatment + sex + (treatment|isofemale), data  = heat, family = "binomial")
summary(mod3)
anova(mod3, mod2)
dotplot(ranef(mod3))$isofemale

# plot(fixef(mod3), fixef(mod2))
# abline(0,1)
# require(ggplot2)
# ggplot(data = NULL, aes(x = fixef(mod3), y = fixef(mod2)))+
#   geom_point() +
#   geom_abline(slope = 1, intercept = 0, lwd = 2, col = 'red') +
#   theme_bw()

## there's almost a perfect 1:1 relationship between the fixed effects for the model with and without the fixed*random interaction. So it doesn't add a very significant difference in the value.
```

